### 2016-06-23 paper/icfp16-cameraready/evaluation.tex
try placing all traces in sec:advantage-traces in a single figure
actually, let's group all of the graphs
leave the traces inline
done

\paragraph{Benchmarks}
We answer the first two questions on two sets of ill-typed programs,
\ie\ programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages (CSE 130) course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs.
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
polymorphic data types, and expression evaluators.

@@

\begin{figure*}[ht]
\centering
\begin{minipage}{\linewidth}
\centering
\includegraphics[width=0.49\linewidth]{coverage.png}
% \end{minipage}
% \begin{minipage}{\linewidth}
\includegraphics[width=0.49\linewidth]{distrib.png}
\end{minipage}
% \vspace{-8ex}
\caption{Results of our coverage testing and the distribution of test
  outcomes. Our random search successfully finds witnesses for 79--85\% of
  the programs in under one second, improving to 87\% in under
  10 seconds. In both datasets we detect actual type errors about 82\%
  of the time, unbound variables or constructors 3--4\% of the time, and
  diverging loops 2--3\% of the time. For the remaining 11--12\% of the
  programs we are unable to provide any useful feedback.  }
\label{fig:results-witness}
\end{figure*}

@@

For each of the ill-typed programs for which we could
find a witness, we measure the complexity of the generated
trace according to two metrics.

% \paragraph{Metrics} Thus, our two metrics are:
% size of the full trace,
% \ie the number of small-step reductions, and the size of the jump-compressed
% version of the trace.
%
\begin{enumerate}
\item \emphbf{Single-step Metric} The size of the trace after expanding
  all of the single-step edges from the witness to the stuck term, and
  % This can be thought of as a worst-case
  % complexity, \ie ``How big is the fully-expanded trace?''
\item \emphbf{Jump-compressed Metric} The size of the jump-compressed trace.
\end{enumerate}

% \item \ES{others?}
%
\begin{figure*}[ht]
\centering
\includegraphics[width=0.49\linewidth]{trace_size_step.png}
\includegraphics[width=0.49\linewidth]{trace_size_jump.png}
\caption{Complexity of the generated traces. 81\% of the combined traces
  have a jump complexity of at most 10, with an average complexity of 7
  and a median of 5.}
\label{fig:results-complexity}
\end{figure*}
%

\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 31 for the
\ucsdbench\ dataset (35 for the \uwbench\ dataset) with a maximum of
2,745 (986 for \uwbench) and a median of 17 (also 17 for \uwbench).
%
The average number of jumps per trace is 7 (also 7 for \uwbench) with a
maximium of 353 (185 for \uwbench) and a median of 4 (also 4 for
\uwbench).
%
In both datasets 80\% or more traces have at most 10 jumps.
%

\subsection{Witness Utility}
\input{gallery2}

\subsubsection{Quantitative}
\label{sec:user-study}
% Finally, to test the explanatory power of our jump-compressed traces, we
% ran a user study (IRB \#2014009900) at the University of Virginia (UVA).
% %
% We included four problems in an exam in the Spring session of UVA's
% undergraduate Programming Languages course (CS 4501).
% %
% We presented the 60 students in the course with ill-typed \ocaml\
% programs and asked them to
% %
% (1) \emph{explain} the type error, and
% %
% (2) \emph{fix} the type error.
% %
% For each problem the student was given the ill-typed program and
% either \ocaml's error message or \toolname's jump-compressed trace.
%
We assigned four problems to the ($n=60$) students in the course: the
@sumList@, @digitsOfInt@, and @wwhile@ programs from
\S~\ref{sec:advantage-traces}, as well as the following @append@ program
%
\begin{ecode}
  let append x l =
    match x with
    | []   -> l
    | h::t -> h :: t :: l
\end{ecode}
%
which triggers an occurs-check error on line 4.
%
For each problem the students were given the ill-typed program and
either \ocaml's error message or \toolname's jump-compressed trace.
%
Due to the nature of an in-class exam, not every student answered every
question; we received between 13 and 28 (out of a possible 30) responses
for each problem-tool pair.

@@

\paragraph{Threats to Validity}
\subparagraph{Construct}
%
Measuring understanding is a difficult task; we used the correctness of
the student's explanation of, and fix for, the type error as a proxy for
her understanding, but it is possible that other metrics would produce
different results.

@@

\paragraph{Results}
%
Figure~\ref{fig:results-user-study} summarizes a single annotator's
results, which show that students given \toolname's jump-compressed
trace were consistently more likely to correctly explain
and fix the type error than those given \ocaml's error message.
%
Across each problem the \toolname responses were marked correct
$10-30\%$ more often than the \ocaml responses, which suggests that
the students who had access to \toolname's traces had a better
understanding of the type errors.
%
The measured kappa values were $\kappa = 0.72$ for the explanations and
$\kappa = 0.83$ for the fixes; while there is no formal notion for what
consititutes strong agreement~\cite{Krippendorff2012-wd}, kappa values
above $0.60$ are often called ``substantial''
agreement~\cite{Landis1977-ey}.
%
\begin{figure*}[ht]
\centering
\includegraphics[width=0.49\linewidth]{user-study-reason.png}
\includegraphics[width=0.49\linewidth]{user-study-fix.png}
\caption{A classification of students' explanations and fixes for type
  errors, given either \ocaml's error % message
  or \toolname's
  jump-compressed trace. The students given \toolname's jump-compressed
  trace consistently scored better ($\ge 10\%$) than those given \ocaml's type
  error.}
\label{fig:results-user-study}
\end{figure*}
%

>>>>

\paragraph{Benchmarks}
We answer the first two questions on two sets of ill-typed programs,
\ie\ programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages (CSE 130) course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs.
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
polymorphic and algebraic data types. % and expression evaluators.

@@

\begin{figure*}[htp]
\centering
\begin{minipage}{\linewidth}
\centering
\includegraphics[width=0.45\linewidth]{coverage.png}
% \end{minipage}
% \begin{minipage}{\linewidth}
\includegraphics[width=0.45\linewidth]{distrib.png}
\end{minipage}
% \vspace{-8ex}
\caption{Results of our coverage testing and the distribution of test
  outcomes. Our random search successfully finds witnesses for 79--85\% of
  the programs in under one second, improving to 87\% in under
  10 seconds. In both datasets we detect actual type errors about 82\%
  of the time, unbound variables or constructors 3--4\% of the time, and
  diverging loops 2--3\% of the time. For the remaining 11--12\% of the
  programs we are unable to provide any useful feedback.  }
\label{fig:results-witness}
\end{figure*}

@@

For each of the ill-typed programs for which we could
find a witness, we measure the complexity of the generated
trace using two metrics.

% \paragraph{Metrics} Thus, our two metrics are:
% size of the full trace,
% \ie the number of small-step reductions, and the size of the jump-compressed
% version of the trace.
%
\begin{enumerate}
\item \emphbf{Single-step:} The size of the trace after expanding
  all of the single-step edges from the witness to the stuck term, and
  % This can be thought of as a worst-case
  % complexity, \ie ``How big is the fully-expanded trace?''
\item \emphbf{Jump-compressed:} The size of the jump-compressed trace.
\end{enumerate}

% \item \ES{others?}
%
\begin{figure*}[htp]
\centering
\includegraphics[width=0.45\linewidth]{trace_size_step.png}
\includegraphics[width=0.45\linewidth]{trace_size_jump.png}
\caption{Complexity of the generated traces. 81\% of the combined traces
  have a jump complexity of at most 10, with an average complexity of 7
  and a median of 5.}
\label{fig:results-complexity}
\end{figure*}
%
\begin{figure*}[htp]
\centering
\includegraphics[width=0.45\linewidth]{user-study-reason.png}
\includegraphics[width=0.45\linewidth]{user-study-fix.png}
\caption{A classification of students' explanations and fixes for type
  errors, given either \ocaml's error % message
  or \toolname's
  jump-compressed trace. The students given \toolname's jump-compressed
  trace consistently scored better ($\ge 10\%$) than those given \ocaml's type
  error.}
\label{fig:results-user-study}
\end{figure*}

\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 31 for the
\ucsdbench\ dataset (35 for the \uwbench\ dataset) with a maximum of
2,745 (986 for \uwbench) and a median of 17 (also 17 for \uwbench).
%
The average number of jumps per trace is 7 (also 7 for \uwbench) with a
maximium of 353 (185 for \uwbench) and a median of 4 (also 4 for
\uwbench).
%
In both datasets 80\% or more traces have at most 10 jumps.

\subsection{Qualitative Evaluation of Witness Utility}\label{sec:advantage-traces}

Next, we present a \emph{qualitative} evaluation that compares
the explanations provided by \toolname's dynamic witnesses with
the static reports produced by the \ocaml\ compiler and \sherrloc,
a state-of-the-art fault localization approach~\cite{Zhang2014-lv}.
%
In particular, we illustrate, using a series of examples drawn
from student programs in the \ucsdbench\ dataset, how \toolname's
jump-compressed traces can get to the heart of the error. Our approach
%
highlights the conflicting values that cause the program to get
stuck, rather that blaming a single one,
%
shows the steps necessary to reach the stuck state, and
%
does not assume that a function is correct just because it type-checks.
%
For each example we will present
(1)~the code,
(2)~the error message returned \ocaml,
(3)~the error locations returned by \ocaml\ (underlined) and \sherrloc\ (in bold),\footnote{When the locations from \ocaml\ and \sherrloc\ overlap,
we just underline the relevant code.}
and (4)~the jump-compressed trace produced by \toolname.

% \begin{figure*}[ht]
% \centering
% \begin{minipage}{0.49\linewidth}
% \centering

% \begin{figure*}[htp]
%   \centering
%   \includegraphics[height=125px]{sqsum.png}
%   \includegraphics[height=125px]{sumlist.png}
%   \includegraphics[height=150px]{digitsOfInt.png}
%   \includegraphics[height=125px]{wwhile.png}
%   \caption{(Left to right) Jump-compressed traces showing how
%     \texttt{sqsum}, \texttt{sumList}, \texttt{digitsOfInt}, and
%     \texttt{wwhile} go wrong in \S~\ref{sec:advantage-traces}.}
%   \label{fig:traces}
% \end{figure*}

\paragraph{Example: Recursion with Bad Operator}
The recursive function @sqsum@ should square each
element of the input list and then compute the sum
of the result.
%
\begin{ecode}
  let rec sqsum xs = match xs with
    | [] -> 0
    | h::t -> __(sqsum t)__ @ (h * h)
\end{ecode}
%
Unfortunately the student has used the list-append
operator |@| instead of \texttt{+} to compute the sum.
%
Both \ocaml\ and \sherrloc\ blame the \emph{wrong location},
the recursive call @sqsum t@, with the message
%
\begin{verbatim}
  This expression has type
    int
  but an expression was expected of type
    'a list
\end{verbatim}
%
\toolname\ produces a trace showing how the evaluation of
@sqsum [1]@ gets stuck.
%
\begin{center}
  \includegraphics[height=125px]{sqsum.png}
\end{center}
%
The trace highlights the entire stuck term
(not just the recursive call), emphasizing
the \emph{conflict} between @int@ and @list@
rather than assuming one or the other is correct.

\paragraph{Example: Recursion with Bad Base Case}
%
The function @sumList@ should add up
the elements of its input list.
%
\begin{ecode}
  let rec sumList xs = match xs with
    | []    -> ==[]==
    | y::ys -> y + __sumList ys__
\end{ecode}
%
Unfortunately, in the base case, it returns @[]@
instead of @0@.
%
\sherrloc\ blames the base case, and \ocaml\
assumes the base case is correct and blames
the \emph{recursive call} on line 3:
%
\begin{verbatim}
  This expression has type
    'a list
  but an expression was expected of type
    int
\end{verbatim}
%
Both of the above are parts of the full story, which
is summarized by \toolname's trace showing
how @sumList [1; 2]@ gets stuck at @2 + []@.
%
\begin{center}
  \includegraphics[height=125px]{sumlist.png}
\end{center}
%
The trace clarifies (via the third step)
that the @[]@ is the result of the recursive call
@sumList []@, and shows how it is incompatible with
the subsequent \texttt{+} operation.

%% ES: append is actually a bit problematic as we don't find the nice
%% append [1] [2] witness. instead we could find something like
%% append [_] [], but it's not as clear IMO
% Our next example is the @append@ function, which should concatenate the
% two input lists.
% %
% \begin{ecode}
% let append xs ys = match xs with
%   | []   -> ys
%   | h::t -> h :: __t__ :: ys
% \end{ecode}
% %
% The student has forgotten to make a recursive call to @append@, and
% instead tries to cons the tail @t@ directly onto the second list @ys@.
% Consing @h@ back onto the result causes \ocaml to attempt to construct
% the infinite type @'a = 'a list@, triggering an \emph{occurs-check}
% error.
% %
% \begin{verbatim}
% Error: This expression has type
%          'a list
%        but an expression was expected of type
%          'a
%        The type variable 'a occurs inside 'a list
% \end{verbatim}
% %
% %
% \begin{center}
%   \includegraphics[height=75px]{append.png}
% \end{center}

%\pagebreak
\paragraph{Example: Bad Helper Function that Type-Checks}
%
The function @digitsOfInt@ should return a list of
the digits of the input integer.
%
\begin{ecode}
  let append x xs =
    match xs with
    | [] -> [x]
    | _  -> x :: xs

  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      append (==digitsOfInt (n / 10)==)
             [__n mod 10__]
\end{ecode}
%
%\pagebreak
Unfortunately, the student's @append@ function \emph{conses} an element
onto a list instead of appending two lists.
%
Though incorrect, @append@ still type-checks and thus \ocaml and
\sherrloc blame the \emph{use-site} on line 10.
%
\begin{verbatim}
  This expression has type
    int
  but an expression was expected of type
    'a list
\end{verbatim}
%
In contrast, \toolname makes no assumptions about @append@
yielding a trace that illustrates the true error on line 4, by
highlighting the conflict in consing a list onto a list of integers.
%
\begin{center}
  \includegraphics[height=160px]{digitsOfInt.png}
\end{center}
%

\paragraph{Example: Higher-Order Functions}
%
The higher-order function @wwhile@ is supposed
to emulate a traditional while-loop. It takes
a function @f@ and repeatedly calls @f@ on the
first element of its output pair, starting with
the initial @b@, till the second element is @false@.
%
\begin{ecode}
  let rec wwhile (f,b) =
    match f with
    | (z, false) -> z
    | (z, true)  -> wwhile (f, z)

  let f x =
    let xx = x * x in
    (xx, (xx < 100))

  let _ = wwhile (__f__, 2)
\end{ecode}
%
The student has forgotten to \emph{apply} @f@ at all on line 2,
and just matches it directly against a pair.
This faulty @wwhile@ definition nevertheless typechecks,
and is assumed to be correct by both \ocaml\ and \sherrloc\
which blame the use-site on line 10.
%
\begin{verbatim}
  This expression has type
    int -> int * bool
  but an expression was expected of type
    'a * bool
\end{verbatim}
%
\toolname\ synthesizes a trace that draws the eye to the
true error: the @match@ expression on line 2, and highlights
the conflict in matching a function against a pair pattern.
%
\begin{center}
  \includegraphics[height=135px]{wwhile.png}
\end{center}
%
By highlighting conflicting values, \ie\ the source and sink
of the problem, and not making assumption about function correctness, \toolname\
focusses the user's attention on the piece of code that is
actually relevant to the error.

\subsection{Quantitative Evaluation of Witness Utility}
\label{sec:user-study}
% Finally, to test the explanatory power of our jump-compressed traces, we
% ran a user study (IRB \#2014009900) at the University of Virginia (UVA).
% %
% We included four problems in an exam in the Spring session of UVA's
% undergraduate Programming Languages course (CS 4501).
% %
% We presented the 60 students in the course with ill-typed \ocaml\
% programs and asked them to
% %
% (1) \emph{explain} the type error, and
% %
% (2) \emph{fix} the type error.
% %
% For each problem the student was given the ill-typed program and
% either \ocaml's error message or \toolname's jump-compressed trace.
%
We assigned four problems to the ($n=60$) students in the course: the
@sumList@, @digitsOfInt@, and @wwhile@ programs from
\S~\ref{sec:advantage-traces}, as well as the following @append@ program
%
\begin{ecode}
  let append x l =
    match x with
    | []   -> l
    | h::t -> h :: t :: l
\end{ecode}
%
which triggers an occurs-check error on line 4.
%
For each problem the students were given the ill-typed program and
either \ocaml's error message or \toolname's jump-compressed trace.
%
Due to the nature of an in-class exam, not every student answered every
question; we received between 13 and 28 (out of a possible 30) responses
for each problem-tool pair.

@@

\paragraph{Threats to Validity}
Measuring understanding is a difficult task; the following summarize
the threats to the validity of our results.

\subparagraph{Construct}
%
We used the correctness of the student's explanation of, and fix for,
the type error as a proxy for her understanding, but it is possible
that other metrics would produce different results.

@@

\paragraph{Results}
%
Figure~\ref{fig:results-user-study} summarizes a single annotator's
results, which show that students given \toolname's jump-compressed
trace were consistently more likely to correctly explain
and fix the type error than those given \ocaml's error message.
%
Across each problem the \toolname responses were marked correct
$10-30\%$ more often than the \ocaml responses, which suggests that
the students who had access to \toolname's traces had a better
understanding of the type errors.
%
The measured kappa values were $\kappa = 0.72$ for the explanations and
$\kappa = 0.83$ for the fixes; while there is no formal notion for what
consititutes strong agreement~\cite{Krippendorff2012-wd}, kappa values
above $0.60$ are often called ``substantial''
agreement~\cite{Landis1977-ey}.
%
%

### 2016-06-23 paper/icfp16-cameraready/introduction.tex
done

\paragraph{1. Finding Witnesses}
Our first contribution is an algorithm for searching for
\emph{witnesses} to type errors, \ie\ inputs that cause a
program to go wrong~(\S~\ref{sec:searching-witness}).
%
This problem is surprisingly tricky when we cannot rely on
static type information.
%
In particular, we must avoid the trap of \emph{spurious} inputs
that cause irrelevant problems that would be avoided by picking
values of a different, relevant type.
%
We solve this problem by developing a novel operational semantics
that combines evaluation and type inference.
%
We execute the program with \emph{holes} --- values whose type is
unknown --- as the inputs.
% that corresponds to executing the program with \emph{holes} ---
% values whose type is unknown.
%
A hole remains abstract until the evaluation context tells us what type
it must have, for example the parameters to an addition operation must
both be integers.
%
Our semantics \emph{conservatively} instantiates holes with concrete
values, dynamically inferring the type of the input
until the program goes wrong.
%
We prove that our procedure synthesizes \emph{general witnesses}, which
means, intuitively, that if a witness is found for a given ill-typed
function, then, \emph{for all} (inhabited) input types, there exist
values that can make the function go wrong.

>>>>

\paragraph{1. Finding Witnesses}
Our first contribution is an algorithm for searching for
\emph{witnesses} to type errors, \ie\ inputs that cause a
program to go wrong~(\S~\ref{sec:searching-witness}).
%
This problem is tricky when we cannot rely on
static type information, as we must avoid the
trap of \emph{spurious} inputs that cause
irrelevant problems that would be avoided
by picking values of a different, relevant type.
%
We solve this problem by developing a novel
operational semantics that combines evaluation
and type inference.
%
We execute the program with \emph{holes} --- values whose type is
unknown --- as the inputs.
%
A hole remains abstract until the evaluation
context tells us what type it must have, for
example the parameters to an addition operation
must both be integers.
%
Our semantics conservatively instantiates holes
with concrete values, dynamically inferring the
type of the input until the program goes wrong.
%
We prove that our procedure synthesizes \emph{general}
witnesses, which means, intuitively, that if a witness
is found for a given ill-typed function, then, \emph{for all}
(inhabited) input types, there exist values that can make
the function go wrong.

### 2016-06-23 paper/icfp16-cameraready/related.tex
shrink
done

\section{Related Work}
\label{sec:related-work}
In this section we connect our work to related efforts on type errors,
testing, and program exploration.

@@

\paragraph{Localizing and Repairing Type Errors}
\label{sec:diagnosis-repair}
It is well known that unification-based type inference procedures can
produce poor error messages, and in particular, can misidentify the
\emph{source} of the type error.
%
Thus, many groups have explored techniques to pinpoint the true source
of the error and recommend fixes.

% \subparagraph{Localization}
The traditional Damas-Milner type inference
algorithm~\cite{Damas1982-uw} reports the first program location where a
type mismatch is discovered (subject to the traversal
strategy~\cite{Lee1998-ys}).
%
As a result the error can be reported far away from its
source~\cite{McAdam1998-ub} without enough information to guide the
user.
%
Type-error slicing~\cite{Haack2003-vc,Schilling2011-yf,Rahli2015-tt,Sagonas2013-bf,Gast2004-zd,Neubauer2003-xv}
%treats type inference as a constraint-satisfaction problem and
recognizes this flaw and instead produces a slice of the program
containing \emph{all} program locations that are connected to the type
error.
%
%% \citet{Neubauer2003-xv} present a decidable type system based on
%% discriminative sum types, in which all terms are typeable and type
%% derivations contain all type errors in a program. They then use the
%% typing derivation to slice out the parts of the expression related to
%% each error.
%
Though the program slice must contain the source of the error, it can
suffer from the opposite problem of providing \emph{too much}
information, motivating recent work in ranking the candidate locations.
%
Zhang~\etal~\cite{Zhang2014-lv,Zhang2015-yu} present an algorithm for
identifying the most likely culprit using Bayesian reasoning.
%
\citet{Pavlinovic2014-mr,Pavlinovic2015-kh} translate the %error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.

@@

\paragraph{Improving Error Messages}
The content and quality of the error messages themselves has also been
studied extensively.
%
\citet{Marceau2011-ok,Marceau2011-cy} study the effectiveness of error
messages in novice environments and present suggestions for improving
their quality and consistency.
%
\citet{Hage2006-hc} identify a variety of general heuristics to improve
the quality of type error messages, based on their teaching experience.
%
\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
\citet{Serrano2016-oo} provide methods for library authors to specialize
type errors with domain-specific knowledge.
%
The difference with our work is more pronounced here as we do not
attempt to improve the quality of the error message, instead we search
for a witness to the error and explain it with the resulting execution
trace.

@@

\paragraph{Testing}
\label{sec:testing}
\nanomaly is at its heart a test generator, and builds on a rich line of
work.
%
Our use of holes to represent unknown values is inspired by the work of
Runciman, Naylor, and Lindblad~\cite{Runciman2008-ka,Naylor2007-mi,Lindblad2007-oy},
%
who use lazy evaluation to drastically reduce the search space for
exhaustive test generation, by grouping together equivalent inputs by
the set of values they force. An exhaustive search is complete (up to
the depth bound), if a witness exists it will be found, but due to the
exponential blowup in the search space the depth bound can be quite
limited without advanced grouping and filtering techniques.
%
Our search is not exhaustive; instead we use random generation to fill
in holes on demand.
%
Random test generation~\cite{Claessen2000-lj,Csallner2004-bf,Pacheco2007-at}
%
is by its nature incomplete, but is able to check larger inputs than
exhaustive testing as a result.

@@

@@

\paragraph{Program Exploration}
Perera \etal~\cite{Perera2012-dy} present a tracing semantics
for functional programs that tags values with their provenance, enabling
a form of backwards program slicing from a final value to the sequence
of reductions that produced it. Notably, they allow the user to supply a
\emph{partial value} --- containing holes --- and present a partial slice,
containing only those steps that affected the the partial value.
% This
% system is designed to answer questions of the form ``Where did this
% value come from?'' and thus is focused on backward exploration. 
Perera~\etal\ focus on backward exploration; in contrast, our
visualization supports forward \emph{and} backward exploration, though
our backward steps are more limited.
%
Specifically, we do not support selecting a value and inserting the
intermediate terms that preceded it while ignoring unrelated computation
steps. %; this would be interesting future work.

>>>>

\section{Related Work}
\label{sec:related-work}
% In this section we connect our work to related efforts on type errors,
% testing, and program exploration.

@@

\paragraph{Localizing and Repairing Type Errors}
\label{sec:diagnosis-repair}
% It is well known that
% unification-based type inference procedures can
% produce poor error messages, and in particular, can misidentify the
% \emph{source} of the type error.
%
Many groups have explored techniques to pinpoint the true source
of errors reported by static type checkers.
% unification-based type inference procedures.
%
% \subparagraph{Localization}
The traditional Damas-Milner type inference algorithm~\cite{Damas1982-uw}
reports the first program location where a type mismatch is discovered
(subject to the traversal strategy~\cite{Lee1998-ys}).
%
As a result the error can be reported far away from its
source~\cite{McAdam1998-ub} without enough information to guide the
user.
%
Type-error slicing~\cite{Haack2003-vc,Schilling2011-yf,Rahli2015-tt,Sagonas2013-bf,Gast2004-zd,Neubauer2003-xv}
%treats type inference as a constraint-satisfaction problem and
recognizes this flaw and instead produces a slice of the program
containing \emph{all} program locations that are connected to the type
error.
%
%% \citet{Neubauer2003-xv} present a decidable type system based on
%% discriminative sum types, in which all terms are typeable and type
%% derivations contain all type errors in a program. They then use the
%% typing derivation to slice out the parts of the expression related to
%% each error.
%
Though the program slice must contain the source of the error, it can
suffer from the opposite problem of providing \emph{too much}
information, motivating recent work in ranking the candidate locations.
%
Zhang~\etal~\cite{Zhang2014-lv,Zhang2015-yu} present an algorithm for
identifying the most likely culprit using Bayesian reasoning.
%
\citet{Pavlinovic2014-mr,Pavlinovic2015-kh} translate the %error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.

@@

\paragraph{Improving Error Messages}
%
The content and quality of the error messages themselves has also been
studied extensively.
%
\citet{Marceau2011-ok,Marceau2011-cy} study the effectiveness of error
messages in novice environments and present suggestions for improving
their quality and consistency.
%
\citet{Hage2006-hc} identify a variety of general heuristics to improve
the quality of type error messages, based on their teaching experience.
%
\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
\citet{Serrano2016-oo} provide methods for library authors to specialize
type errors with domain-specific knowledge.
%
The difference with our work is more pronounced here as we do not
attempt to improve the quality of the error message, instead we search
for a witness to the error and explain it with the resulting execution
trace.
%

@@

\paragraph{Testing}\label{sec:testing}
%
\nanomaly is at its heart a test generator, and as such,
builds on a rich line of work.
%
Our use of holes to represent unknown values is inspired by the work of
Runciman, Naylor, and Lindblad~\cite{Runciman2008-ka,Naylor2007-mi,Lindblad2007-oy},
%
who use lazy evaluation to drastically reduce the search space for
exhaustive test generation, by grouping together equivalent inputs by
the set of values they force. An exhaustive search is complete (up to
the depth bound), if a witness exists it will be found, but due to the
exponential blowup in the search space the depth bound can be quite
limited without advanced grouping and filtering techniques.
%
Our search is not exhaustive; instead we use random generation to fill
in holes on demand.
%
Random test generation~\cite{Claessen2000-lj,Csallner2004-bf,Pacheco2007-at}
%
is by its nature incomplete, but is able to check larger inputs than
exhaustive testing as a result.

@@

A function's type is a theorem about the function's behavior.
Thus, \toolname's witnesses can be viewed as \emph{counterexamples},
thereby connecting it to work on using test generation to find
counterexamples prior to starting a proof proof~\cite{ACL2Testing, Seidel15}.

@@

\paragraph{Program Exploration}

\cite{Flanagan96} describes a static debugger for Scheme, which helps
the programmer interactively visualize problematic source-sink flows
corresponding soft-typing errors. The debugger allows the user to explore
an abstract reduction graph computed from a static value set analysis of
the program. In contrast, \toolname generates witnesses and allows the user
to explore the resulting dynamic execution.
%
Perera \etal~\cite{Perera2012-dy} present a tracing semantics
for functional programs that tags values with their provenance, enabling
a form of backwards program slicing from a final value to the sequence
of reductions that produced it. Notably, they allow the user to supply a
\emph{partial value} --- containing holes --- and present a partial slice,
containing only those steps that affected the the partial value.
% This
% system is designed to answer questions of the form ``Where did this
% value come from?'' and thus is focused on backward exploration.
Perera~\etal\ focus on backward exploration; in contrast, our
visualization supports forward \emph{and} backward exploration, though
our backward steps are more limited.
%
Specifically, we do not support selecting a value and inserting the
intermediate terms that preceded it while ignoring unrelated computation
steps. %; this would be interesting future work.
